Galois Characterization of Function Fields of \\ Transcendence Degree $1$ over Number Fields

### 2017-01-18 Master Thesis/synopsis.tex
Solve TeX warnings
Clarify proof of 4.5

\begin{lemma}\label{2.4}
Let $K$ be not necessarily finitely generated and $\ell\neq 2$ be a prime. The following map, induced by the restrictions, is injective:
\[ \mathrm{H}^3(K,\mathbb{Z}/\ell\mathbb{Z})\to\prod_{w\in |\operatorname{Spv}(K)| }\mathrm{H}^3(K_w,\mathbb{Z}/\ell\mathbb{Z}) \]
\end{lemma}

@@

\section{Global Correspondence mod $n$}

@@

\begin{proof}
First we assume that $k/\mathbb{Q}$ is a Galois extension and $X$ has a $k$-rational point $P_0$. We can consider $k'$ as a subfield of $k$ via: 
\[\begin{tikzcd}
k'\ar[r] & k(P'_0)\ar[r, "\varphi_{P_0}"', "\cong"] & k(P_0) \ar[r, "\cong"] & k
\end{tikzcd}\]
For all $P\in X_0$ and $P'=\sigma(P)\in X'_0$ we also get an embedding:
\[ \begin{tikzcd}
k \ar[r] & k(P) \ar[r, "\varphi_P^{-1}"', "\cong"] & k(P')
\end{tikzcd} \]
Let us consider the morphism $f:X'\times_{k'} k\to X'$. We want to see that $f$ splits completely. For $P'\in X'_0$ we have as topological spaces:
\[ f^{-1}(P') = \operatorname{Spec} k(P')\times_{k'} k \]
Any embedding $k\to \overline{k(P')}$ maps $k$ into $k(P')$ because $k/\mathbb{Q}$ is Galois. Thus we have:
\[ k(P')\otimes_{k'} k\cong \prod_{\operatorname{Hom}_{k'}(k, k(P'))} k(P') \]
Now we can see that $f$ splits completely indeed. By Lemma \ref{4.1} we get $k\cong k'$, hence $[k:\mathbb{Q}]= [k':\mathbb{Q}]$.

@@

\begin{proof}
\begin{enumerate}[(i)]
\item By Corollary \ref{mordell-weil-corollary}, the Picard group of $X$ is finitely generated. Let $D_1,\ldots, D_r$ be Weil divisors of $X$ such that the associated line bundles generate $\operatorname{Pic}(X)$ and similarly let $D'_1,\ldots,D'_s$ generate $\operatorname{Pic}(X')$. Clearly any finite $S\subset X_0$ such that
\[ \bigcup_{i=1}^r \operatorname{supp}D_i\subset S\quad\text{and}\quad \bigcup_{j=1}^s\operatorname{supp}D'_j\subset S' \]
is admissible. Now let $V\subset X$ be an open subscheme. By the approximation theorem, we can choose the $D_i$ and the $D'_j$ in $V_0$ and $\sigma (V_0)$ respectively.
\item Observe that the following sequence is exact:
\[ \begin{tikzcd}
0 \ar[r] & K^S \ar[r]& K^\times \ar[r, "\operatorname{div}"]& \displaystyle \bigoplus_{P\in X_0\setminus S}\mathbb{Z} \ar[r]& 0
\end{tikzcd} \]
Since the right module is $\mathbb{Z}$-flat, tensoring with $\mathbb{Z}/n\mathbb{Z}$ yields another exact sequence and we can consider $K^S/n$ as a subgroup of $K^\times/n$. Then:
\[ K^S/n = \bigcap_{P\in X_0\setminus X}U_P/n \]
The analogous result holds for $K'^{S'}/n$. Thus the assertion follows from Lemma \ref{4.3} (i).\qedhere
\end{enumerate}
\end{proof}

>>>>

\begin{lemma}[\textit{Main Lemma}]\label{2.4}
Let $K$ be not necessarily finitely generated and $\ell\neq 2$ be a prime. The following map, induced by the restrictions, is injective:
\[ \mathrm{H}^3(K,\mathbb{Z}/\ell\mathbb{Z})\to\prod_{w\in |\operatorname{Spv}(K)| }\mathrm{H}^3(K_w,\mathbb{Z}/\ell\mathbb{Z}) \]
\end{lemma}

@@

\section{Global Correspondence mod \texorpdfstring{$n$}{n}}

@@

\begin{proof}
First we assume that $k/\mathbb{Q}$ is a Galois extension and $X$ has a $k$-rational point $P_0$. We can consider $k'$ as a subfield of $k$ via: 
\[\begin{tikzcd}
k'\ar[r] & k(P'_0)\ar[r, "\varphi_{P_0}"', "\cong"] & k(P_0) \ar[r, "\cong"] & k
\end{tikzcd}\]
For all $P\in X_0$ and $P'=\sigma(P)\in X'_0$ we also get an embedding:
\[ \begin{tikzcd}
k \ar[r] & k(P) \ar[r, "\varphi_P^{-1}"', "\cong"] & k(P')
\end{tikzcd} \]
Let us consider the morphism $f:X'\times_{k'} k\to X'$. We want to see that $f$ splits completely. For $P'\in X'_0$ we have as topological spaces:
\[ f^{-1}(P') = \operatorname{Spec} k(P')\times_{k'} k \]
Any embedding $k\to \overline{k(P')}$ maps $k$ into $k(P')$ because $k/\mathbb{Q}$ is Galois. Thus we have:
\[ k(P')\otimes_{k'} k\cong \prod_{\operatorname{Hom}_{k'}(k, k(P'))} k(P') \]
Now we can see that $f$ splits completely. By Lemma \ref{4.1} we get $[k:k'] = 1$, hence $[k:\mathbb{Q}]= [k':\mathbb{Q}]$.

@@

\begin{proof}
\begin{enumerate}[(i)]
\item By Corollary \ref{mordell-weil-corollary}, the Picard group of $X$ is finitely generated. Let $D_1,\ldots, D_r$ be Weil divisors of $X$ such that the associated line bundles generate $\operatorname{Pic}(X)$ and similarly let $D'_1,\ldots,D'_s$ generate $\operatorname{Pic}(X')$. Clearly any finite $S\subset X_0$ such that
\[ \bigcup_{i=1}^r \operatorname{supp}D_i\subset S\quad\text{and}\quad \bigcup_{j=1}^s\operatorname{supp}D'_j\subset S' \]
is admissible. Now let $V\subset X$ be an open subscheme. By the approximation theorem, we can choose the $D_i$ and the $D'_j$ in $V_0$ and $\sigma (V_0)$ respectively.
\item Observe that the following sequence is exact:
\[ \begin{tikzcd}
1 \ar[r] & K^S \ar[r]& K^\times \ar[r, "\operatorname{div}"]& \displaystyle \bigoplus_{P\in X_0\setminus S}\mathbb{Z} \ar[r]& \operatorname{Pic}(X\setminus S) = 0
\end{tikzcd} \]
Since $\bigoplus\mathbb{Z}$ is $\mathbb{Z}$-flat, tensoring with $\mathbb{Z}/n\mathbb{Z}$ yields another exact sequence and we can consider $K^S/n$ as a subgroup of $K^\times/n$. Then:
\[ K^S/n = \bigcap_{P\in X_0\setminus X}U_P/n \]
The analogous result holds for $K'^{S'}/n$. Thus the assertion follows from Lemma \ref{4.3} (i).\qedhere
\end{enumerate}
\end{proof}
